---
title: "Practical 3"
author: "Analysing data"
---

<!-- To knit this .Rmd do the following -->

```{r setup_how_to_knit, include=F, eval=F}
remotes::install_github("mivalek/teachR")
teachR::make.sheet("[path to file]/ad_practical03.Rmd", "and", toc_depth = 3, tasks_to_headings = T)
```


## Today

- Read in and check data
- Basic descriptive statistics
- Including tables in papers/reports
- Write-up: Participants

## Guided part

```{r setup_data, echo = F}
set.seed(1)
library(tidyverse)
gensex <- read_csv("https://mivalek.github.io/adata/gen_sex_q.csv")
```

First of all, let's make sure we're all set up:

- Log into Canvas and go to this week's practical
- Download the "prac_03_basic_desc.Rmd" file into your analysing_data project folder
- Open RStudio

### Read in data

Data from the form you filled in in the previous lecture are here: https://mivalek.github.io/adata/gen_sex_q.csv

Anything online is just some sort of file. In this case, it is a "comma-separated values" file, or CSV. We *could* download it to our folder and then read it into RStudio. However, we can also read it directly from the URL like this:

```{r, eval=F}
read_csv("https://mivalek.github.io/adata/gen_sex_q.csv")
```

**Time to think**

Why might it be preferable to read data in from the URL?

Why might it be better to save data on your computer?

The `read_csv()` function comes from the `tidyverse` package. What do we need to do before we can run the command above?

What will the command above do? Is that what we want? If not, what do we need to add to the line?

Let's read the data into RStudio so that we can start working with them.

```{r, eval=F}
library(tidyverse)
gensex <- read_csv("https://mivalek.github.io/adata/gen_sex_q.csv")
```

### First glance

Now that we have our data in RStudio, we should familiarise ourselves with it. Best place to start is often looking at the structure of the data set.

```{r}
gensex %>% str()
```

Here we can see if the variables are coded as numbers (`num`), character strings "`char`", or logical (`logi`). The first step in data audit/cleaning is to make sure these so called "classes" (`numeric`, `character`, `logical`) are what we would expect given our variables and, if not, to understand why.

Looking at the ouput of the command above, everything is looking good, except for `Age`. `Age` is supposed to be a numeric variable, but `R` read it as a character string. To understand why this happened, we need to look at what values were entered as age by our participants. We can do that by grouping our data by `Age` and then taking the tally of different ages:

```{r}
gensex %>%
  group_by(Age) %>%
  tally()
```

### Data cleaning

Hmm.. Looks like some participants didn't just enter the number, but instead wrote "18 years" and "19 years old". Because of these two data points, the variable cannot be understood as numbers. To convert `Age` into a numeric variable, we first need to replace these values with 18 and 19, respectively. We can write a pipeline that *mutates* (changes) the value of the `Age` variable *recoding* the two rows as numbers:

```{r}
gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19"))
```

OK, something happened, but it's sort of difficult to see what and whether we did what we wanted. But because of the flexibility of the pipelines, we can just take this resulting tibble and group it by `Age` and take the tally again, just like we did before:

```{r}
gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19")) %>%
  group_by(Age) %>%
  tally()
```

This looks good: WHereas before we had 69 participants with age 18 and 113 with age 19, now we have 70 and 114, respectively.

We're not done yet, though! The variable is still coded as `character` so we need to change it into `numeric` by adding the `as.numeric()` function, inside the `mutate()` section of our pipeline:

```{r}
gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19"),
         Age = as.numeric(Age))
```

We can check that this did the job by pulling out the `Age` variable from the dataset and looking at its class:

```{r}
gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19"),
         Age = as.numeric(Age)) %>%
  pull(Age) %>%
  class()
```

Success!

<div class = "warn">
Notice that the `"18"` and `"19"` are in quotes. This is important. Remember that `Age` is currently a `character` variable so if we are changing some of its values, the replacement values must also be `character` values. Putting in simply `18` and `19` will result in unwanted behaviour:

```{r}
gensex %>%
  mutate(Age = recode(Age, "18 years" = 18, "19 years old" = 19),
         Age = as.numeric(Age))
```
</div>

\ 

Don't forget that pipelines on their own **do not modify the original data**. To save our changes into `gensex`, we need to reassign the ouput of the pipeline back to itself:

```{r}
gensex <- gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19"),
         Age = as.numeric(Age))
```

\ 

One more thing: It's good to set categorical variables to be factors instead of character vectors. Factors are handy for plotting and some statistics.

```{r}
gensex <- gensex %>%
  mutate(Gender = factor(Gender))
```

We can double-check all is now as its supposed to be:

```{r}
gensex %>% str()
```

\ 

Now that basic housekeeping is done, let's look at the roughest summary of our data.

```{r}
gensex %>% summary()
```

As we can see, we have some missing values in the `Age` and `Gender` variables. That's OK, missing data happen.

However, because 

Notice, however, a few oddities. Firstly, all variables contain at least 19 missing values. It is possible that there are participants who started the questionnaire but never actually replied to any questions. We want to delete these "attempts".

#### Removing rows with only `NA`s

Let's think about how we can tell `R` to remove those rows of our dataset that only have missing values. The following steps do the trick:

1. Test each data point for whether or not it is an `NA`
2. Add up the results *by row*
3. Remove rows for which this added up number is equal to the number of variables of interest

OK, let's take it one step at a time.

To test each data point for missingness, we can use the `is.na()` function. The ouput of this function is not a `tibble` but we can convert it into one using `as.tibble()`:

```{r}
is.na(gensex) %>% as.tibble()
```

As you can see, we now get only the values `TRUE` and `FALSE` in place of our data. We can use the fact that computers teat `TRUE` as 1 and `FALSE` as 0 to calculate the number of missing values per each row. To do that, we only need to take the row sums using the aptly names `rowSums()` function:

```{r}
rowSums(is.na(gensex))
```

Because we only have 12 variables in our questionnaire (plus the Duration column), any row that has 12 missing values doesn't include any data. Thus, we can filter only those rows that have fewer than 12 `NA`s. Before we do that, lets count how many rows we will be removeing as this is something we should mention in our write-up:

```{r}
rowSums(is.na(gensex)) == 12
all_missing <- sum(rowSums(is.na(gensex)) == 12)
all_missing
```

OK, so we're removing `r all_na` rows. Let's do it using `filter()`.

```{r}
gensex <- gensex %>% filter(rowSums(is.na(gensex)) < 12)
```

<!-- \  -->

<!-- However, `id` should be a unique code so it's good to check if there are no duplicates. The `duplicated()` function takes a vector and returns a vector of `TRUE`s and `FLASE`s of equal length. We expect all values to be `FALSE` (*e.g.,* `id`s are unique). -->

<!-- ```{r} -->
<!-- duplicated(c(1, 2, 5)) # unique -> all FALSE -->
<!-- duplicated(c(1, 2, 2)) # not unique -> some TRUE -->
<!-- ``` -->

<!-- Let's create a column `id_check` and ask for `summary()` again -->

<!-- ```{r} -->
<!-- gensex %>% -->
<!--   mutate(id_check = duplicated(id)) %>% -->
<!--   summary() -->
<!-- ``` -->

<!-- Cool, there are only `FALSE`s which means no duplicates! -->

<!-- *Note. *We don't really need the `id_check` column permanently in our data set so we don't need to modify it by doing `gensex <- gensex %>% mutate(...`. -->

### Basic descriptive stats

With all of this out of the way, we can prepare tibbles of relevant summary statistics. First of all, age:

```{r}
age_desc <- gensex %>%
  summarise(age_mean = mean(Age),
            sd_age = sd(Age),
            min_age = min(Age),
            max_age = max(Age),
            na_age = sum(is.na(Age)))
age_desc
```

Why are we getting weird results?

What do we need to do to get what we want?

```{r}
age_desc <- gensex %>%
  summarise(mean = mean(Age, na.rm = T),
            sd = sd(Age, na.rm = T),
            min = min(Age, na.rm = T),
            max = max(Age, na.rm = T),
            na = sum(is.na(Age)))
age_desc
```

We only ever need to report these kinds of numbers to 2 decimal places so let's round things up:

```{r}
age_desc <- age_desc %>% 
  modify(round, 2)
age_desc
```

We can do all of the above in a single *pipeline*:

```{r}
age_desc <- gensex %>%
  summarise(mean = mean(Age, na.rm = T),
            sd = sd(Age, na.rm = T),
            min = min(Age, na.rm = T),
            max = max(Age, na.rm = T),
            na = sum(is.na(Age))) %>%
  modify(round, 2)
age_desc
```

This `age_desc` tibble now enables us to report descriptive statistics in text using inline code. For instance, we can just type <code>&#96;r age_desc %>% pull(mean)&#96;</code> in R Markdown and when the file gets generated, it will read `r age_desc %>% pull(mean)`. That's useful, isn't it!

\ 

Let's create another tibble of descriptives, this time with age breakdown by Gender:

```{r}
gen_age_desc <- gensex %>%
  group_by(Gender) %>%
  summarise(n = n(),
            perc = n()/nrow(gensex) * 100,
            mean_age = mean(Age, na.rm = T),
            sd_age = sd(Age, na.rm = T)) %>%
  modify_if(is.numeric, round, 2)
gen_age_desc
```

The command returns the following message:

"Factor `Gender` contains implicit NA, consider using `forcats::fct_explicit_na`"

This is not technically a problem, we can just have a row for `NA` in the tibble. However, using the `fct_explicit_na()` function suggested by the warning creates an additional level `(Missing)` in the `Gender` variable, instead of the `NA`s. We can use the function inside of `mutate()`:

```{r}
gen_age_desc <- gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>%
  group_by(Gender) %>%
  summarise(n = n(),
            perc = n()/nrow(gensex) * 100,
            mean_age = mean(Age, na.rm = T),
            sd_age = sd(Age, na.rm = T)) %>%
  modify_if(is.numeric, round, 2)
gen_age_desc
```

I think this looks better and there are also minor advantages of having explicitly labelled missing data when it comes to plotting. More on that later...

\ 

This second descriptive tibble is also quite handy as we can take it and easily convert in into a nice table using the `kable()` function from package `knitr`:

```{r}
library(knitr)
gen_age_desc %>% kable()
```

Granted, the table isn't formatted perfectly... Let's do it then!

There are basically only three things we want to edit

1. It would be good to capitalise the first letters for each gender category
1. The table header could look better
1. The table should have a caption

With respect to the first point, we could laboriously change the first column of `gen_age_desc`. Fortunately, there is a function in the `stringr` package that turns a string into a sentence, *i.e.,* capitalises the first letter:

```{r}
library(stringr)
str_to_sentence("some string we want to edit")
```

We can use this function in `mutate()` to change the first column of our `gen_age_desc` tibble:

```{r}
gen_age_desc %>%
  mutate(Gender = str_to_sentence(Gender))
```

The remaining two edits can be done using arguments to `kable()`, namely `col.names=` and `caption=`. We can use R Markdown within these arguments to get things like subscripts or italics:

```{r}
gen_age_desc %>%
  mutate(Gender = str_to_sentence(Gender)) %>%
  kable(col.names = c("Gender", "*N*", "%", "*M*~age~", "*SD*~age~"),
        caption = "Table 1 *Descriptive statistics by Gender*")
```

\ 

That's pretty neat, don't you think?

\ 

<div class="warn">
Tables in documents generated from your own R markdown files will look slightly different. That is because we are using our own custom theme for these sheets.
</div>


### Write-up

Now that we are familiar with our data, we can start writing up. Obviously, we don't really have any results yet but we can at least describe our sample. This is normally done in the *Participants* subsection of the *Method* section. On top of that, we can give a description of what data were collected and how in the *Materials* and *Procedure* subsections, respectively. Here, both would be very short so we can collapse them into a single subsection:

#### Method

##### Participants

Data from `r nrow(gensex)` first year Psychology students at the University of Sussex (*M*~age~ = `r age_desc %>% pull(mean)`, *SD*~age~ = `r age_desc %>% pull(sd)`) were collected anonymously as part of the Analysing Data module.

##### Materials and procedure

Participants filled in a questionnaire hosted via Qualtrics.
The form was accessed using a URL given in lecture.
<!-- Each participant was given a unique ID code they then entered into the form. -->
The questionnaire contained ten Likert-type items on topics relating to gender and sexuality. Data on participants' own age and gender were also collected. The age breakdown by gender is given in Table 1.

```{r, echo=F}
gen_age_desc %>%
  mutate(Gender = str_to_sentence(Gender)) %>%
  kable(col.names = c("Gender", "*N*", "%", "*M*~age~", "*SD*~age~"),
        caption = "Table 1 *Descriptive statistics by Gender*")
```

\ 

<div class="solText">
<title>Points to notice</title>

Read through our mock write-up again and notice the following points:

  - Concision
      - Make it short and sweet. Above, we only needed one sentence to tell the reader who our participants were, how many there were in our sample, what the basic descriptives of their age were, how they were recruited and where and how it all happened.
  - Right amount of detail
      - You want to report all the info about your sample that we did. We *could* also have reported the range of ages but it's not crucial.
      - In *Material and procedure* we need to give the reader enough information so that they can get a good idea of what exactly it is that we did.
  - You can refer to tables and figures if it's helpful
      
These are all good guidelines for writing up the *Method* section.
</div>

\ 

That's all we have to cover this week. Over to you now!

## Worksheet

In the remaining time you will practice what you have just learnt. To make things interesting, let's use the data from the Gender and sexuality questionnaire we filled in in the lecture. You can find the data at https://...

`r task()`First of all, if you haven't done it yet, download the "01_basic_desc.Rmd" file from canvas and open it in RStudio.

Use the R Markdown file to complete the following tasks.

`r task()`In the `packages` code chunk, write the code to load all the packages you will need to complete this practical: `tidyverse`, `knitr`, `stringr`, and `cowplot` should be enough.

`r task()`In the `read-in` code chunk, write the code to read the data into RStudio.

`r task()`In the `inspect` chunk, write code that does the following:

`r subtask()`Make sure the classes of your variables match the code book: Ordinal variables can be coded as numbers but categorical variables should be factors.

*Hint. *A basic summary will do.

<!-- `r subtask()`Check if all ID codes are unique. -->

`r task()`Take a minute to look at the summary of the data set. Notice the means and medians of the numeric variables. Think about what the distributions of the variables might look like.

`r task()`In the `descriptives` code chunk, write code that creates:

`r subtask()`A tibble of descriptive statistics (mean, standard deviation, minimum, maximum) for the `age` variable.

`r subtask()`A tibble with *N*s, %s, and age (mean and *SD*) breakdown by categories of the `gender` variable.

`r task()`In the `plots` code chunk, write code tat generates the following basic plots:

`r subtask()`A histogram of `age`:

```{r, echo = F}
# gensex %>%
#   ggplot(aes(x = age)) +
#   geom_histogram(bins = 8, color = "black", fill = "grey") +
#   theme_cowplot()
```

`r subtask()`A barchart of `gender`:

**ADD PLOT**

`r subtask()`A boxplot of `age` by `gender`:

**ADD PLOT**

`r task()`If you haven't formatted your plots yet, change the code in the chunk to generate prettier plots. If you want to, go mad with colours! [Here are some colour names you can use.](http://sape.inf.usi.ch/sites/default/files/ggplot2-colour-names.png)

`r task()`Complete the **Write-up** section of the .Rmd file.

`r task()`Edit the code in the `table_1` code chunk, giving it your tibble with age breakdown by gender, to create a nice formatted table in your document.

`r task()`Knit (generate) the final document from your R markdown file and rejoice in its beauty.

\ 

That's all for this week. Well done!

\ 

\ 


