---
title: "Lecture 1: Introduction"
subtitle: "Analysing Data, Spring 2020"
author: "Dr Jennifer Mankin"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(tidyverse)
```

## Welcome to Analysing Data!

* Overview of the Module
  + Summary of topics and lectures
  + Format and assessments
* Data and Science
  + Data for this term: gender and sexuality
  + What makes good science?
* Welcome BACK to R
  + What you can do so far
  + What you will be able to do
  + Tips for success

## Overview: Topics and Lecturers

Find a complete list of all topics for each week [on Canvas](https://canvas.sussex.ac.uk/courses/9242/assignments/syllabus)!

* Part 1: Fundamentals of Statistical Inference
  + Lecturer: [Dr Milan Valasek](http://www.sussex.ac.uk/profiles/489434/dropin)
  + Sampling, distributions, estimates, hypothesis testing
* Part 2: Basic Statistical Models
  + Lecturer: [Dr Jennifer Mankin](http://www.sussex.ac.uk/profiles/351056/dropin)
  + Measures of association, the linear model
* Part 3: Next Steps
  + Lecturer: [Dr Rebecca Atkinson](http://www.sussex.ac.uk/profiles/363651/dropin)
  + Expanding the linear model, the big picture

## Overview: Format

* Lectures
  + One hour each week
  + Concepts, demonstrations, ideas, etc.
  
* Practicals
  + Two hours each week
  + Hands-on application of statistical concepts in R

## Overview: Attendance

You are required to attend your scheduled practical session. We will **not** arrange ad-hoc swaps for practicals. 

If you have a consistent timetabling clash, you must email [the UG Office](psychology-ug@sussex.ac.uk) to change your timetable.

## Overview: Assessments

Assessment             | Length             | Weight | Due
-----------------------| ------------------ | ------ | ----------
In-Class Quizzes       | Best 8 of 10       | 20%    | Every week
Lab Report             | 2000 words         | 20%    | Week 9
Research Participation | 16 credits/8 hours | 10%    | Week 12
Unseen Exam            | 2 hours/50 MCQs    | 50%    | A2 Exams

## Assessments: Quizzes

* As last semester, each week you will have a tutorial to complete before your practical class
* Every week, you will have a **marked quiz** on the lecture and tutorial
  * 7 questions (multiple choice, fill in the blank, simple calculations, etc.)
  * 7 minutes to complete (9 minutes with extra time)
  * 10 quizzes total, in every week **except this week**
    * Practice quiz in all practicals this week!
* Your final quiz mark is the mean of your best 8 out of 10 total quiz scores
  
## Assessments: Lab Report

* Last term you wrote a report on one of two research questions:
  * Red: Does colour influence attractiveness?
  * Green: Does social status drive environmentally conscious decisions?
  
* This term, you will write a report on experimental data testing one of these two research questions
  * You can freely choose the same or a different study as last term

## Assessments: Lab Report

* You will submit an RMarkdown document that includes:
  + Brief Participants and Design sections
  + A results section presenting your analysis, with a graph
  + A Discussion section evaluating the implications, limitations, future directions etc.

* See [Lab Report Information and Resources](https://canvas.sussex.ac.uk/courses/9242/pages/Lab%20Report%20Information%20and%20Resources?titleize=0) on Canvas

## Assessments: Professional Log

* As last term, you must complete 16 credits (4 hours) of research participation via [SONA](https://sussexpsychology.sona-systems.com/Default.aspx?ReturnUrl=%2f)
  + At least 25% (4 credits/1 hour) of these credits must be gained through in-person studies

* See [SONA information and FAQs](https://canvas.sussex.ac.uk/courses/9242/pages/research-participation-module-credits-scheme) on Canvas

## Assessments: Unseen Exam

* In the A2 Spring assessment period, you will have a two-hour multiple-choice exam
* More detail on this near the end of term!

# Data and Science

* Data for this term: Gender and sexuality
* What makes good science?

# Gender and Sexuality

* Both are often represented as simple (?) categories
* Important to understand!
    + Huge impact on how people perceive, treat, think about themselves and others
* "The living world is a **continuum** in each and every one of its aspects." - Kinsey et al. (1948, p. 639)
* This semester we will look at these constructs in some depth

## Operationalising Gender

* Gender very often represented as a **categorical variable** in psychology
  + Usually binary (two categories): "male" and "female" (sometimes with "other")
  + Also confounded with [biological sex](https://www.scientificamerican.com/article/beyond-xx-and-xy-the-extraordinary-complexity-of-sex-determination/)
* Is this binary really representative of what gender is really like?
* What **is** gender really like? How can we measure it?
  + In short: complicated (*very* complicated)

## Operationalising Sexuality

* Like gender, sexual orientation is often presented as categories
  + More complex here: *which* categories?
  + "Traditional": heterosexual, homosexual, bisexual, asexual
  + What about: pansexual, demisexual, andro/gynesexual, etc.?
* Additionally complicated by **romantic** orientation and other factors
* It seems there are the same number of categories as there are humans!

## Operationalising Sexuality: the Kinsey Scale

Rating | Description
------ | ------------------------------------------------------------------
  0    | Exclusively heterosexual
  1	   | Predominantly heterosexual, only incidentally homosexual
  2    | Predominantly heterosexual, but more than incidentally homosexual
  3    | Equally heterosexual and homosexual
  4    | Predominantly homosexual, but more than incidentally heterosexual
  5    | Predominantly homosexual, only incidentally heterosexual
  6    | Exclusively homosexual
  X    | No socio-sexual contacts or reactions

## Gender and Sexuality Questionnaire

* We have posted questionnaire on Canvas about gender and sexuality
* In the practicals, we will look at patterns in this data
  + Are these constructs best measured as categories? As scales? Which ones?
  + How do the responses relate to each other?
* Ethical considerations:
  + Participation is **strictly voluntary**
    - Absolutely no impact on you, your mark, or anything else
  + All responses will be entirely anonymous
  + Please complete **by Friday the 31st**!

# What makes good science?

* Validity
  + Are we actually measuring what we think we're measuring?
* Reliability
  + Do we get consistent/reproducible results using this measurement?
* Reproducibility and the Scientific Method
  + How can we be confident that what we believe is the truth?

## Validity

* Are we actually measuring what we think we're measuring?
* Fundamental distinction between a construct and its measurement
  + e.g. Academic ability vs marks earned
* A more complicated and difficult question than you might think!
  + We **only** have access to the measurement - not to the construct

## Reliability

* Do we get consistent/reproducible results using this measurement?
  + Using a meter stick to measure a person's height - very reliable!

## Reliability and Validity Together

* Consider general intelligence and IQ tests
  + Reliability: do you score (about) the same number each time you take an IQ test?
  + Validity: If you have an IQ of 110 and your friend has an IQ of 105, are you 4.5% cleverer than they are?

## Replication and the Scientific Method

* How can we be confident that what we believe is the truth?
* You can be *confident* - but you can **never be sure**!
* Science never proves - only gives evidence

### The Scientific Method

* Make an observation or propose a theory
* Create a disproveable hypothesis
* Design a test of the hypothesis and predict what will happen
* Run the test and obtain data
* Interpret the data
* Reach a conclusion and generate more hypotheses

### Lessons from the Scientific Methods

* Things that are too good to be true usually are!
* Beware of claims that fit exactly with what you expected, or wanted
* Don't believe **anything** without evidence
  + Always look at the evidence for yourself

### The Replication Crisis

* Take two minutes to talk to the person next to you:
  + What is the "replication crisis" in Psychology?
  + Why has it happened?
  + What can we do about it?

### The Replication Crisis

* What is the "replication crisis" in Psychology?
  + Findings from famous, well-known studies do not replicate
    - i.e. when someone else tries to run the same study (usually with a larger and more diverse sample), they do not get the same results as the original
    - Example: [Power posing](https://www.youtube.com/watch?v=r7dWsJ-mEyI) (see [this blog post](https://statmodeling.stat.columbia.edu/2017/10/18/beyond-power-pose-using-replication-failures-better-understanding-data-collection-analysis-better-science/) for more)
    - Example: [Ego depletion](http://www.slate.com/articles/health_and_science/cover_story/2016/03/ego_depletion_an_influential_theory_in_psychology_may_have_just_been_debunked.html?via=gdpr-consent) (see [this article](https://www.theatlantic.com/science/archive/2018/11/psychologys-replication-crisis-real/576223/) for more)
    
### The Replication Crisis

* Why has it happened?
* A very good question with many possible answers and no definitive ones!
* Some contributing factors:
  + Poor understanding of methods and statistics
  + Pressure to publish novel and exciting results
  + Poor research design
  + Consistent use of very specialised, ungeneralisable (e.g. [WEIRD](https://psmag.com/social-justice/joe-henrich-weird-ultimatum-game-shaking-up-psychology-economics-53135)) participants
  
### The Replication Crisis

* What can we do about it? So glad you asked!
  + Design thoughtful, thorough, well-researched, meaningful experiments
  + Pre-register (plan ahead) study and analysis
  + Report and publish **all** results, regardless of the results
  + Make data and analyses openly available
    - Often by writing clear, helpful R scripts!

# Welcome Back to R!

## What you can do so far in R

* Use R and RStudio
* Create projects and Markdown files
* Read in data and view tibbles
* Use functions to choose, group, and summarise data
* Use the pipe operator (`%>%`) with tidy data and `tidyverse` functions
* Create and interpret frequency distributions with `ggplot2`
* Format and knit Markdown files with images, links, etc.

## Lost in tRanslation

* R is complicated but working hard at it **will pay off**
  + Transferable skills for (many!!!) future jobs
  + Programming/data management skills are in [high demand](https://royalsociety.org/news/2019/05/data-science-skills-shortages/)

* Think of learning R as learning to speak another language
  * You can already say some basic phrases in R
  * This term we will keep building your grammar and vocabulary

## Translate this!

How can you read this code?

```{r, echo = TRUE, eval = FALSE}
data_summary <- data %>% 
  group_by(gender) %>% 
  summarise(n = n())
```

* A: Make a new object, `data_summary`, by taking `data`, grouping it by `group`, and then summarising the number of n's in the data
* B: Make a new object, `data_summary`, by taking `data`, grouping it by `gender`, and then summarising the number of cases
* C: Make a new object, `data`, by summarising the total number of cases and then grouping by `gender`

## What you will be able to do

* Create tables graphs of data
* Describe estimates and measures of variance
* Perform statistical tests for measuring associations and predicting results
* Produce clean, professional summary reports, including graphs and statistical results

# Tips for Success

1. Practice practice practice!
2. Get help!
3. Don't give up!

## Tip 1: Practice practice practice!

* The more you use R, the easier it will be
* Try lots of different tasks with the data you have
  * Quiz yourself!
* Try out other [online learning resources for R](https://canvas.sussex.ac.uk/courses/9242/pages/r-resources) to get more practice

## Tip 2: Get help!

* Work through tutorials and practicals with friends on the module and problem-solve together
* Use the resources on the [Reading List](https://canvas.sussex.ac.uk/courses/9242/external_tools/328) for solutions and ideas
* Search for help online - there are countless videos, blogs, Twitter feeds, etc. to help you!
* Come to lecturer drop-in hours and/or R Helpdesk [(times posted on Canvas)](https://canvas.sussex.ac.uk/courses/9242) to ask questions

## Tip 3: Don't get discouraged!

* Think of R as a logic puzzle to solve
* Set yourself (or look for) tasks and games
* You will get better with practice if you keep at it!

# Looking Ahead

* For this week, you will have a **practice quiz** in your practical session
  + You can take the practice quiz as many times as you like
  + Next week's marked quiz will be on the same questions!
* Please complete the Gender and Sexuality Questionnaire on Canvas
  + We will use this data for the practicals, so the more people who do it the better!
* See you in the practicals!
