---
title: "Measures of Association"
subtitle: "Analysing Data Week 5"
author: "Dr Jennifer Mankin"
date: "[Click for handout](./lecture_5_assoc.html)"
---

```{r setup, include=FALSE, message=F, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)
library(kableExtra)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(weights)
library(adata)
```

```{r report.p, include = F}
report.p <- function(x){
  ifelse(x > .001, paste0(" = ", rd(x, 3)), "< .001")
}
```

##  Overview of the Next Weeks
 
```{r, echo=F, results='asis'}
cat("
<style>
:root {
--theme-col: #03a7a7;
--hover-col: #A70303;
}
</style>
")
```

```{r, revealjs-setup, include=F}
knitr::opts_chunk$set(comment=NULL, collapse=T, strip.white=F,
                      warning=F, message=F,
                      fig.height = 3.5,
                      fig.width = 5)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
hook_inline = knitr::knit_hooks$get('inline')
knitr::knit_hooks$set(
  inline = function(x) {
    res = hook_inline(x)
    if (is.numeric(x)) prettyNum(format(x, scientific=FALSE), big.mark=',') else res
  })
```

* Today: Measures of Association
* Week 6: Comparing Two Means
* Week 7: Lab Report Writing
* Week 8: The Linear Model

##  Overview of Today's Topics

* Review of Important Ideas
  + Continuous vs Categorical Data
  + Distributions, test statistics, and NHST
* Tests of Association
  + What can we learn from these tests?
  + Continuous Data: Correlation coefficient *r*
  + Frequency Data: Chi-squared test *&chi;*^2^

##  Review of Important Ideas

* Data can be measured in many ways
  + Continuous, categorial, ratio, ordinal...
* The type of data can have a big impact on your analysis and interpretion of results

##  Continuous Data
+ A matter of *degree*; answers the question "How much?"
  - Typically a score or measurement
  - Examples: a quiz marks, height in cm, response latency
  - If it makes sense to have a *mean*, you likely have continuous data

##  Categorical Data
+ A matter of *membership*; answers the question "Which one?"
  - Typically a group or label
  - Examples: practical group, alive vs. dead, control vs experimental condition
  - Membership in any category is binary (either in or out)

##  Distributions, Test Statistics, and NHST

* Everything from the past few weeks we will now put into action!
* For each statistical analysis, we will have the same ingredients:
  + Data
  + A test statistic
  + The distribution of that test statistic
  + The probability *p* of the value of the test statistic we have under the null hypothesis H~0~

##  Reminder about NHST

* For all of our tests, we assume that **the null hypothesis is true** unless we have evidence to suggest that this is unlikely
  + "Evidence" such as a test statistic that is sufficiently unlikely to occur under H~0~
  + In that case, we reject the null and accept the alternative hypothesis H~1~
* In psychology, "sufficiently unlikely" is typically defined as less than 5% probability (*p* < .05)
  + Nothing magical about this number!
  
##  Overall Reminder

* As scientistis we want to believe true things about the world, and disbelieve false things
* NHST (*p*-values), and statistics more generally, are systems to help us make decisions about whether, and to what degree, we believe something is true

##  Measures of Association

* Essential questions: are two variables related to each other, or unrelated (independent)? How can we quantify this relationship?
  + Choose a test statistic that best captures the relationship you want to model
* Today's tests:
  + Correlation (*r*)
  + Chi-Squared (*&chi;*^2^)

##  Correlation

* Quantifies the **degree** and **direction** of a numeric relationship
* Typically used with two (or more) continuous variables
  + Can be used when one is categorical!
* Today's example: Gender and Sexuality data from the questionnaire

##  Correlation: Visualisation

```{r fem_masc_plot, fig.width=8, fig.height=5}
gensex <- read_csv("../../data/gensex_clean.csv")

p1 <- gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>% 
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1)) +
  geom_point(position = "jitter", size = 2, alpha = .4) +
  labs(x = "Femininity", y = "Masculinity", title = "Plot of ratings of femininity vs masculinity") +
  theme_cowplot()
p1
```

##  Correlation

* The plot shows that people who gave high ratings for femininity tended to give low ratings for masculinity, and vice versa
* Things we might like to know:
  + How strong is this relationship?
  + Should we believe that it's real (ie representative of people/first-year psychology students in general?)

##  Correlation: How strong?

* We can quantify the **strength** and **direction** of the relationship between femininity and masculinity with Pearson's correlation coefficient *r*
* Strength: absolute value of *r* between 0 and 1
  + 0: no relationship at all
  + 1: perfect relationship
* Direction: whether the value of *r* is positive or negative
  + Positive: as one variable increases, the other tends to increase
  + Negative: as one variable increases, the other tends to decrease
* So, values range from -1 (perfect negative) through 0 (no relationship) to 1 (perfect positive)

##  Correlation: Let's Try It!

```{r gensex_cor, echo = T}
gensex %>% 
  select(Gender_fem_1, Gender_masc_1) %>% 
  cor(method = "pearson")
```
```{r gensex_cor_rep}
gensex_r <- gensex %>% 
  select(Gender_fem_1, Gender_masc_1) %>% 
  cor(method = "pearson") %>%
  pluck(2)
```
* So, our correlation coefficient *r* is `r gensex_r %>% round(2)`
  + The negative sign (-) means as femininity increases, masculinity tends to decrease (and vice versa)
  + We saw this on the plot as a downward trend left to right
  + The absolute value of `r gensex_r %>% round(2) %>% abs()` is **very** strong - quite close to 1!

##  Correlation: Is It Real?

* We now have our data and our test statistic *r* (`r gensex_r %>% round(2)`)
* We can now ask likely we are to get a value of `r gensex_r %>% round(2)` if in fact femininity and masculinity have a true *r* of 0
  + i.e. the null hypothesis is in fact true
  + We will use the standard significance level of < .05 in this case

##  Correlation: Is It Real?

```{r, gensex_cor_test, echo = T}
fm.r <- cor.test(gensex$Gender_fem_1, gensex$Gender_masc_1, alternative = "two.sided", method = "pearson")
fm.r
```

We can report this as: "There was a significant correlation between femininity and masculinity, *r*(`r fm.r$parameter`) = `r gensex_r %>% round(2)`, *p* `r fm.r$p.value %>% report.p()`."

##  Correlation: *r* or *t*?

Our output includes *t* and its associated *p*-value:

```{r}
fm.r
```

This is because the probability *p* associated with *r* actually comes from a *t*-distribution with N - 2 degrees of freedom!

##  Correlation: So Far

* We have seen that the correlation coefficient *r* expresses the strength and direction of the relationship between two variables
* We can use this to **compare** the strength of different relationships

##  Correlation: Comparing *r*s

```{r fem_masc_plot_gender, fig.width=8, fig.height=5}
gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>% 
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1, colour = Gender)) +
  geom_point(position = "jitter", size = 2, alpha = .4) +
  labs(x = "Femininity", y = "Masculinity", title = "Ratings of self-reported femininity vs masculinity by gender identity") +
  theme_cowplot()
```

##  Correlation: Interpretation

* For both female-ID and male-ID people, there is a fairly strong negative relationship
* We can use *r* to quantify whether this relationship is stronger for female-ID or male-ID people
  + Simply compare the values of *r* for females and males separately

##  Correlation: Comparing values

```{r fem_vs_male_corr}
male.gs <- filter(gensex, Gender == "Male")
fem.gs <- filter(gensex, Gender == "Female")

male.cor <- cor.test(male.gs$Gender_fem_1, male.gs$Gender_masc_1, alternative = "two.sided", method = "pearson")

fem.cor <- cor.test(fem.gs$Gender_fem_1, fem.gs$Gender_masc_1, alternative = "two.sided", method = "pearson")
```

* For female-ID people, *r*(`r fem.cor$parameter`) = `r fem.cor$estimate %>% round(2)`, *p* `r fem.cor$p.value %>% report.p()`
* For male-ID people, *r*(`r male.cor$parameter`) = `r male.cor$estimate %>% round(2)`, *p* `r male.cor$p.value %>% report.p()`
* So, the relationship between femininity and masculinity is stronger for female-ID people
* WARNING: we should be careful about this, because we had `r gensex %>% filter(Gender == "Female") %>% nrow()` female-ID participants and `r gensex %>% filter(Gender == "Male") %>% nrow()` male-ID participants!
  + How can we check the accuracy of our estimates of *r* for each gender ID?
  
##  Correlation: Comparing values

* For female-ID people, *r* was `r fem.cor$estimate %>% round(2)`, 95% CIs [`r fem.cor$conf.int[1] %>% round(2)`, `r fem.cor$conf.int[2] %>% round(2)`] so +/- `r (fem.cor$estimate - fem.cor$conf.int[1]) %>% round(2)`
* For male-ID people, *r* was `r male.cor$estimate %>% round(2)`, 95% CIs [`r male.cor$conf.int[1] %>% round(2)`, `r male.cor$conf.int[2] %>% round(2)`] so +/- `r (male.cor$estimate - male.cor$conf.int[1]) %>% round(2)`
* Having more female-ID people means our CIs are smaller and estimates are more accurate
  + As these CIs overlap substantially, it's possible that the true value of *r* for male-ID people is not actually different from female-ID people

##  Correlation: **VOCAB ALERT!**

* In common language, "correlated" means "related to in some way, usually causally"
  + Statistics: the (standardised) degree to which two or more variables covary, ie change in relation to each other
* "Correlation" is a technical term!
  + In your reports, do not say two things are "correlated" unless you plan to report *r* as evidence!
  + Instead: variables have a relationship/are related to each other
  
##  Correlation: Summary

* The correlation statistic *r* helps us quantify and compare the strength and direction of relationships between variables
  + How *r* is calculated, relationship to *t*, etc: see reading!
* Correlation **DOES NOT IMPLY CAUSATION!!!!!!!**
* More practice with interpreting *r* with [this fun little game](http://guessthecorrelation.com/)

##  Making Progress

* Correlations: sorted!
  + But we'll come back to them later...
* Next up: Chi-Square (*&chi;*^2^)

##  Chi-Square Test *&chi;*^2^

* Quantifies the relationship between frequencies
* Typically used with two (or more) categorical variables
  + Shows up in **many** contexts!
  + Also used for comparing different models...more in second year!
* Today's example: Simulated data based on the green study, Griskevicius et al. (2010)

##  Green Study: Design

* Research question: Why do people choose to buy "green" products?
  + Conceptual hypothesis: buying "green" products is related to status motives
* Griskevicius et al. (2010) operationalisation:
  + Story condition: read high-status story or status-neutral story
  + Product choice: green or more luxurious non-green product
    - Three products: car, soap, and dishwasher
  + Operational hypothesis: people in the **high-status** condition will choose the **green** products more often than people in the control condition


```{r chisq_tables, include = F}
green_tib <- green_data(seed = 080588)

green_tib <- green_tib %>% 
  mutate(AGE = recode(AGE, "2e" = "2"), 
         AGE = as.numeric(AGE),
         CONDITION = recode(CONDITION, "conrol" = "control"),
         CHOICE = CATEGORY) %>%
  select(-CATEGORY) %>% 
  filter(AGE > 17)

green_sum <- green_tib %>% 
  filter(PRODUCT == "car") %>% 
  select(CONDITION, CHOICE) %>% 
  mutate(CONDITION = recode(CONDITION, "experimental" = "High-Status Story", "control" = "Control Story")) %>%
  group_by(CONDITION) %>% 
  table()

green_chsq <- chisq.test(green_sum)
```

##  Green Study: Data

Here's what the data may look like

* CONDITION: control story or experimental (high-status) story
* CHOICE: product choice, either green or luxury
* We'll just look at the results for one of the three products - the car

```{r chisq_tib}
green_tib
```

##  Green Study: Summary Table

```{r}
green_tab <- green_tib %>% 
  mutate(CONDITION = recode(CONDITION, "experimental" = "High-Status Story", "control" = "Control Story")) %>% 
  filter(PRODUCT == "car") %>% 
  select(CONDITION, CHOICE) %>% 
  group_by(CHOICE) %>% 
  table()

green_tab %>% 
  kable(caption = "Observed Frequencies of Car Product Choices by Story Type",
        col.names = c("Non-Green Luxury Product", "Green Product")) %>% 
  kable_styling()

```


##  Green Study: Graph

```{r chisq_plot, fig.width=8, fig.height= 5}
green_bar <- ggplot(as_tibble(green_tab), aes(x = CONDITION, y = n)) +
  geom_bar(
    aes(fill = CHOICE), 
    stat="identity", position = position_dodge(0.8),
    width = 0.7) +
  labs(title = "Frequency of Car Product Choices by Story Type", x = "Story Condition", y = "Frequency", fill = "Product Choice") +
  scale_y_continuous(limits = c(0, 60)) +
  scale_color_manual(values = c("#3bc53b", "#5c3f78"))+
  scale_fill_manual(values = c("#3bc53b", "#5c3f78"), labels = c("Green Product", "Non-Green Luxury Product"))+
  scale_x_discrete(labels = c("Control Story","High-Status Story"))

green_bar
```

##  Green Study: Analysis

* Are story condition and product choice **independent** of each other?
  + If yes: story condition has no association with product choice
  + If no: story condition is associated with product choice
* The observed frequencies for our sample show some pattern
* How much of a pattern is enough to believe that condition and choice are **not** independent?
  + Compare to what we might expect under the null hypothesis
  + Calculate *&chi;*^2^ to quantify this

##  Chi-Squared: Expected Frequencies

```{r}
green_chsq$expected %>% 
  kable(caption = "Expected Frequencies of Product Choices by Story Condition",
        col.names = c("Non-Green Luxury Product", "Green Product")) %>% 
  kable_styling()
```
* We might expect that there are equal counts in each cell (combination of conditions)
* BUT: To account for unequal numbers of people who chose the green vs non-green products, instead we can calculate the **expected frequencies**
  + That is: how many counts we would expect in each cell, if the null hypothesis is true
  + See An Adventure in Statistics (pgs 434 to 442) for calculations

##  Note on Chi-Squared: Expected Frequencies

* Important assumption: you can only use *&chi;*^2^ if you have expected frequencies greater than 5 in each cell
  + Smaller than this and the test lacks sufficient **power** to detect relationships
* For very small expected frequencies, you may want to use Fisher's Exact Test instead
  + For more on this, see An Adventure in Statistics pgs 442 - 444

##  Chi-Squared Test: Observed vs Expected
* Next, we compare the expected frequencies to our observed frequencies
* The *&chi;*^2^ test statistic represents the (standardised) difference between the observed and expected frequencies
  + The bigger *&chi;*^2^ is, the bigger the difference is between our data and what we expect under the null hypothesis

##  Chi-Squared: Output
  
```{r}
green_chsq
```

+ The calculated value for *&chi;*^2^ (here, "X-squared") is `r green_chsq$statistic %>% round(2)`
+ The degrees of freedom, which determines the *&chi;*^2^ distribution, is `r green_chsq$parameter %>% round(2)` from which we get...
+ The *p*-value, which represents the probability of getting a test statistic this big if the null hypothesis is in fact true
  - Here, *p* is very small (`r green_chsq$p.value %>% report.p()`), so pretty unlikely!
  - We might conclude that the null hypothesis **might not be true**

##  IMPORTANT!!!

* We can only test the significance (probability) of the null hypothesis being true
  + This **does not mean we have evidence for the alternative hypothesis**
* In other words: Our test tells us that it is unlikely that there is no relationship (small *p*-value under the null hypothesis)
  + From that we **infer** that maybe there is a relationship
  + That will only be a sensible thing to infer if we have carefully controlled for all other explanations

##  Green Study: What Can We Conclude?

* We wanted to know whether status motives influence "green" purchases
* Operationalised as whether reading a high-status vs control story was associated with the frequency of green product vs a non-green luxury product choice
* We found that `r (green_chsq$observed[4]/(green_chsq$observed[4]+green_chsq$observed[2]) * 100) %>% round(2)`% of participants who read the high-status story chose a green car over a more luxurious non-green alternative,while only `r (green_chsq$observed[3]/(green_chsq$observed[1]+green_chsq$observed[3]) * 100) %>% round(2)`% of participants who read the control story made the same choice (&chi;^2^(`r green_chsq$parameter`) = `r green_chsq$statistic %>% round(2)`, *p* `r green_chsq$p.value %>% report.p()`).
* So, our results suggest that high-status motives are related to how frequently people choose green products

##  Lab Reports

* You can choose either the red or green study to write your report on
  + See [Lab Report Information on Canvas](https://canvas.sussex.ac.uk/courses/9242/pages/Lab%20Report%20Information%20and%20Resources?titleize=0) for more
* If you choose the green study (Griskevicius et al., 2010), you should use and report the results of *&chi;*^2^
  + Refer to the original paper for some ideas!

##  Analysis for Lab Reports

* Inspect and clean the data
* Produce descriptives of participants (age, gender)
* Report observed frequencies and *&chi;*^2^ results for all three product choices
  + Include at least one figure of the results
  + Will be covered in depth in the next tutorial and practical!
  
##  Questions?
