---
title: 'Lecture 6: Two Means'
author: "Dr Jennifer Mankin"
date: "25/01/2020"
output: html_document
---

```{r}
library(weights)
report.p <- function(x){
  p <- x$p.value
  p.rep <- ifelse(p > .001, paste0(" = ", rd(p, 3)), "< .001")
  print(p.rep)
}
```


# Overview: Two Means

* Study design
  + Defining variables
  + What question are we asking?
* Statistical analysis
  + Student's t-test
  + Under the hood: drawing lines
* Two means in R
  + Welcome to `lm()`!
  + Getting and interpreting results

## Study design

* 

## Statistical analysis



## Two means in R

* Welcome to `lm()`!
* Getting and interpreting results

### Welcome to `lm()`!

* Today's new function is a very important one
  + We will use it for the rest of the term and...
  + It will be very important next year as well!
* What does it do?
  + Creates a **l**inear **m**odel -> `lm()`
  + A statistical model that looks like a line
* How does the linear model work?
  + Much more next week!

### Basic Anatomy of `lm()`

* `lm(outcome ~ predictor, data = data)`
* Translation: 
  + Create a linear model modelling the relationship between
  + An outcome and a predictor
  + Using the data in `data`
* **Try it**: write down the command to predict `attract` from `colour` in our `elliot` data

### Anatomy of `lm()`

`lm(attract ~ colour, data = elliot)`

* Create a linear model that models
* Attractiveness score based on colour group
* Using the data `elliot`

### What will this model actually do?

`attract ~ colour`

* `attract` is a **continuous** variable: each participant's mean score on three rating scales of attractiveness
* `colour` is a **categorical** variable: whether each participant saw the `red` condition or the `white` condition
  + Independent: each participant saw only one of the two

### What will this model actually do?

`attract ~ colour`

1. Group participants by `colour`
2. Calculate the mean `attract` score for each `colour` group
3. Create a linear model that captures the **difference in those means**

## Let's have a look!

```{r, echo = F}
attract_w <- rnorm(2000, mean = 4, sd = 0.5)
attract_r <- rnorm(2000, mean = 7, sd = 0.5)
attract <- c(attract_w,attract_r)
elliot <- tibble(
  colour = factor(rep(c("white", "red"), each = 2000)),
  attract = attract
)

elliot_sum <- elliot %>% 
  group_by(colour) %>% 
  summarise(
    mean = mean(attract),
    sd = sd(attract)
  )

elliot_diff <- elliot_sum$mean[elliot_sum$colour == "red"] - elliot_sum$mean[elliot_sum$colour == "white"]

ggplot(elliot, aes(attract, fill = colour)) + 
   geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity')

# Creating point plot with CIs
ggplot(elliot, aes(x = colour, y = attract, fill = colour)) + 
  stat_summary(fun.data = "mean_cl_normal") # doesn't work atm but might at work!

```

## A closer look

* The mean of the `white` group is `r elliot_sum %>% filter(colour == "white") %>% pull("mean") %>% round(2)`
* The mean of the `red` group is `r elliot_sum %>% filter(colour == "red") %>% pull("mean") %>% round(2)`
* The difference in means is *M*~red~ - *M*~white~ = `r elliot_sum$mean[elliot_sum$colour == "red"] %>% round(2)` - `r elliot_sum$mean[elliot_sum$colour == "white"]` = `r (elliot_sum$mean[elliot_sum$colour == "red"] - elliot_sum$mean[elliot_sum$colour == "white"]) %>% round(2)`

### `attract ~ colour`

* Are these means **different**?
  + Yes: the `colour` groups have different means of `attract`
  + Furthermore: there was, on average, a difference of 3 rating points
  + Translation: women who saw the red border rated the man's attractiveness higher than those who saw the white border
* Are these means different **enough**?
  + How big of a difference is big enough to be meaningful?
  + How big of a difference is big enough to be **significant**?
* Do answer this question, we need to create a model that captures the relationship between `colour` and `attract`

### `attract ~ colour`

Our model will give us two important things:

* An **estimate** of the relationship between `colour` and `attract`
  + Here: the difference between the means in the white and red groups
* The **probability** of finding a difference in means of this size (or greater) occuring if seeing a white or red border has no effect on attractiveness
  + i.e., if the null hypothesis is true

### Model and Results

Let's run the model and see what it is telling us

* This is a lot! For now, focus on "Coefficients"

```{r}
elliot_m <- lm(attract ~ colour, data = elliot)
summary(elliot_m)
```

### Model and Results: **Estimates**

* `Intercept` tells us the predicted value of `attract` when `colour` is 0
  + R picks a baseline (0-coded) category for us
  + Here, `red` is the baseline because it comes first alphabetically
  + This value is familiar: it's the mean of the `red` group

```{r, echo = F}
elliot_m <- lm(attract ~ colour, data = elliot)
summary(elliot_m)
```

### Model and Results: **Estimates**

* `colourwhite` tells us how the mean of `attract` changes when the colour changes from `red` to `white`
  - Here: a change from `red` to `white` corresponds to a drop of about 3 rating points on average
  + This value is familiar: it's the difference between means

```{r, echo = F}
elliot_m <- lm(attract ~ colour, data = elliot)
summary(elliot_m)
```

### Model and Results: **Probabilities**

* New information: the **probability** of finding a difference in means this big (or bigger) if border colour has no effect on attractiveness
  + This is our *p*-value, given under `Pr(>|t|)`
  + We can see that both are smaller than .0000000000000002
  + More helpfully: look at the signif. codes

```{r, echo = F}
elliot_m <- lm(attract ~ colour, data = elliot)
summary(elliot_m)
```

### Model and Results: **Probabilities**

* What does this (extremely small!) probability tell us?
* Let's imagine that if we tested every woman in the world (the whole population), the **real** difference in the means would be 0
  + That is: border colour actually makes no difference
  + In other words, the null hypothesis is true
* If that were true, the probability that we would find a difference in mean attractiveness as big as this one is extremely small!

### Model and Results: **Probabilities**

* The probability that we would get this result if the null hypothesis is true is very small
* We conclude that it's unlikely the null hypothesis is true
* The alternative is that border colour **does** affect attractiveness
* So, we have evidence that the relationship between colour and attractiveness is **significant**

### Interpreting Significance

* In psychology, *p* < .05 is considered "significant"
* There's nothing magical about this number!
  + *p* < .05 is about 1 in 20
  + Changes depends on discipline and precision
* Significance does not "prove" anything!
  + Remember: we only ever gather and evaluate evidence
* Use alongside other ways to quantify differences (e.g. CIs, effect sizes)

### Model and Results: **Probabilities**

* Where does this probability come from anyway?
* Explain the t-distribution here somehow

### The Other Way Round: T-Tests

* Let's try asking for a t.test() instead of an lm()
  + Note: exactly the same specification as with `lm()`

```{r}
elliot_t <- t.test(attract ~ colour, data = elliot)
elliot_t
```

### The Other Way Round: T-Tests

* The `t.test()` output gives us some other useful information
  + Confidence Interval bounds: `r elliot_t$conf.int[1] %>% round(2)` and `r elliot_t$conf.int[2] %>% round(2)`
    - These are CIs around the difference in the means
    - Narrow CIs mean our estimate of the difference is pretty close to the likely population difference
  
```{r, echo = F}
t.test(attract ~ colour, data = elliot)
```

### Reporting a t-test

"An independent samples t-test revealed that participants who saw the white border (*M* = `r elliot_t$estimate[2] %>% round(2)`) rated the target male's attractiveness much lower than those who saw the red border (*M* = `r elliot_t$estimate[1] %>% round(2)`) (*t*(`r elliot_t$parameter %>% round(2)`) = `r elliot_t$statistic %>% round(2)`, *p* `r report.p(elliot_t)`)."

### T-test or linear model?

* It's (essentially) the same thing with different names!
* Typically reported as a "t-test" when this is the only comparison
* Can be reported as an LM when there are more predictors

### Other types of t-tests

* One-sample
* Paired samples