---
title: "Code chunk options"
author: "Analysing Data"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: false
    number_sections: true
    theme: darkly
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = F}
library(tidyverse)
```


## `echo=`

The `echo=` chunk option governs the *rendering* of the `R` code inside a chunk.
By default, it is set to `TRUE` (or `T`).
If it is `TRUE`, the code inside of the chunk gets printed -- or *echoed* -- in the knitted document.
For instance, the code in the following chunk will appear:

```{r }
my_data <- tibble(
  score = rnorm(n = 1000,
                mean = 100,
                sd = 15))
```

`echo=FALSE` hides the code but produces the output

```{r , fig.height=3, fig.width=6, message=F, echo=F, fig.align='center'}
my_data %>%
  ggplot(aes(x = score)) +
  geom_histogram(fill = "#ff450088", colour = "black") +
  theme_classic()
```


## `eval=`

The `eval=` chunk option governs the *evaluation* of the `R` code inside a chunk.
By default, it is set to `TRUE` (or `T`).
If it is set to `FALSE`, the code will appear but will not get run and so any code dependent on objects in this code chunk will fail.

```{r data, eval=F}
my_data2 <- tibble(
  score = rnorm(n = 1000,
                mean = 100,
                sd = 15))
```


Now there is not `my_data2` object and so the plot cannot be generated

```{r histogram, fig.height=3, fig.width=6, message=F, error=T, fig.align='center'}
my_data2 %>%
  ggplot(aes(x = score)) +
  geom_histogram(fill = "#ff450088", colour = "black") +
  theme_classic()
```

## `include=`

The `include=` chunk option cahnges whether or not both code and output should be included.
By default, it is set to `TRUE` (or `T`).
If it is `TRUE`, both code *and* output (plots, tables, *etc*) will be printed in the knitted document.
For instance, here, both code chunks and the plot appear:

<!-- ## `eval=` -->

<!-- The `eval=` chunk option governs the *evaluation* of the `R` code inside a chunk. -->
<!-- By default, it is set to `TRUE` (or `T`). -->
<!-- If it is `TRUE`, the code inside of the chunk will get run by RStudio everytime the document is knitted. -->
<!-- For instance, the code in the following chunk will be evaluated: -->

<!-- ## `echo=` -->

<!-- The `echo=` chunk option governs the *rendering* of the `R` code inside a chunk. -->
<!-- By default, it is set to `TRUE` (or `T`). -->
<!-- If it is `TRUE`, the code inside of the chunk gets printed -- or *echoed* -- in the knitted document. -->
<!-- For instance, the code in the following chunk will appear: -->

```{r, include=T}
my_data <- tibble(
  score = rnorm(n = 1000,
                mean = 100,
                sd = 15))
```


But when `eval=` is set to `FALSE` none of it will be there.

```{r , fig.height=3, fig.width=6, message=F, include=F, fig.align='center'}
my_data %>%
  ggplot(aes(x = score)) +
  geom_histogram(fill = "#ff450088", colour = "black") +
  theme_classic()
```
