---
title: "Analysing Data Practical 1: Setup and Revision"
author: "Your Name"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages(c("WRS2", "GGally", "lm.beta", "kableExtra"))
library(tidyverse)
library(kableExtra)
# Run the following code if tidyverse isn't playing ball on uni computers by deleting the # before each line
# source("http://tinyurl.com/sussexfix")
# sussex.fix()
# sussex::install.sussex("tidyverse")
```

# Part 1: Complete Setup

Make sure you do all of these steps as soon as you arrive:

* Log into OneDrive and Canvas
* Open RStudio via the Software Hub
* Open the Analysing Data project file
* Download and open the RMarkdown and data files for that week
  + Update the Markdown heading (above) with your name
* Run the first (setup) code chunk

## Useful keyboard shortcuts

* In general:
  + Alt + Tab: Switch between open windows
  + Ctrl + a: select all
  + Ctrl + x: cut to clipboard
  + Ctrl + c: copy to clipboard
  + Ctrl + v: paste
  + Ctrl + shift + v: paste as plain text
  + Ctrl + z: undo
  + Ctrl + y: redo
  + Ctrl + s: save
  + Ctrl + f: find in page

For more, see Keyboard Shortcuts [for Windows](https://support.microsoft.com/en-gb/help/12445/windows-keyboard-shortcuts) and [for Macs](https://support.apple.com/en-us/HT201236) 

## More useful keyboard shortcuts

* Within RStudio/RMarkdown
  + Ctrl + Shift + m: %>% 
  + Highlight + (), [], {} etc.: encloses in those brackets
  + Esc: Interrupt currently executing command
  + In the Console
    - Ctrl + 2: Go to console
    - Up/Down: Navigate command history in Console
  + In Markdown
    - Ctrl + Shift + Enter: Run all of current code chunk
    - Ctrl + Enter: Run current line in code chunk

For more, see [Keyboard Shortcuts in RStudio/RMarkdown](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts).

# Part 2: R Practice and Scavenger Hunt

Once you have completed your basic setup, you can start practicing using R again. To do this, you will need to find some data hidden throughout the Analysing Data Canvas site to do some analysis.

Complete the tasks below with the required information. See "Hints" below each set of tasks if you get stuck.

## **Important**: Tasks and Challenges

Make sure you complete the regular tasks first, during the practical session. The tasks marked **Challenge** are designed to help increase your skills in R if you finish early or if you would like to practice more at home.

## Task 1: Data and graphs
### Task 1.1: Fill in the code below 

For all of the values you are asked to find here, use ONLY the Analysing Data Canvas site. Ask a tutor if you get stuck!
 
 * Replace `x` with the maximum word limit for the lab report assessment
 * Replace `m` with the percentage of your total module mark that comes from the weekly quizzes
 * Replace `s` with the number of Research Participation credits you need to obtain via SONA
 
**Hint**: your markdown code should show red x's in the code chunk below - it's trying to warn you that something is missing. Hover your mouse over the red x to see RStudio's suggestion for what's wrong! 
 
```{r}
random_tib <- tibble(
  group = factor(rep(c("ctrl", "exp"), each = 1000)),
  score = rnorm(
    n = x
    mean = m
    sd = s
  )
)
```

### Task 1.2: Plot a histogram of the `score` variable in the `random_tib` dataset

```{r}

```

Task 1.3 **Challenge**: Try the following extra tasks:

* Set the fill hex colour to the last six digits of the University Counselling Service phone number
* Change the X axis label to "Score" and the Y axis label to "Count"
* Group by colour using the "group" variable
* Create side-by-side histograms, one for group = ctrl and one for group = exp

```{r}

```

**Hint**: Use the following tutorials/practicals if you need help:

* [Discovr Summarising Data: Histograms](http://milton-the-cat.rocks/learnr/r/discovr_02/#section-histograms)
* Challenge tasks: [Discovr Visualising Data: ggplot2](http://milton-the-cat.rocks/learnr/r/discovr_05/#section-boxplots-aka-box-whisker-plots)

## Task 2
### Task 2.1: Import the Dataset
### Task 2.2: Create a new summary table
```{r task1}

```

### Task 2.3 **Challenge**: Create a nice table

Use `kable()` and `kable_styling()` to create a nicely formatted table with capitalised column names and the caption "Summary of Length for Mid and High Dose".

```{r}

```

**Hint**: Use the following tutorials/practicals if you need help:

* [Creating factors](http://milton-the-cat.rocks/learnr/r/discovr_01/#section-data-types)
* [Filtering, selecting, and changing variables](http://milton-the-cat.rocks/learnr/r/discovr_01/#section-tibbles)
* [Creating a summary table](http://milton-the-cat.rocks/learnr/r/discovr_02/#section-creating-a-summary-table)
* Using `kable()`: Google it!

## Task 3: Create a report with inline text

### Task 3.1: Create a new code chunk

In a new code chunk, create two new objects, `people` and `blueshirt`, that contain the total number of people you can currently see and the number wearing blue shirts, respectively.

Then, in the following paragraph, use inline code to complete each line by replacing all of the text in square brackets. Try to do this **using only R commands and objects** - do not enter any numbers directly! For example, instead of writing "Today is Tuesday" (when it will not be Tuesday when you read this!), I could write `r Sys.Date() %>% weekdays()`, which should produce the current day of the week whenever you run it. To see how inline text will appear when you knit your document, place your cursor inside the blue code text and push ctrl + enter to run that code.

It is currently [replace with current day]. There are [replace using object] people around me, [replace using calculation] percent of which are wearing blue. 

### Task 3.2 **Challenge**: Complete the report below using the data from Task 1.

Analysis of the dataset `ToothGrowth` revealed that the high-dose group had a mean length of [replace with mean of high dose group] and the mid-dose group had a standard deviation of [replace with sd of mid-dose group]. The overall mean length in the original dataset was [replace with mean of len in tooth_tib].

**Hints**:

* Think about inline code the same way as code in code chunks. You can create multiple commands, linked by `%>%`, to get to the right value.
* Use `pull("name_of_variable")` to extract a particular value from a column in a tibble.
