---
title: "Practical 4 - Visualising data"
author: "Analysing data"
---

<!-- To knit this .Rmd do the following -->

```{r setup_how_to_knit, include=F, eval=F}
remotes::install_github("mivalek/teachR")
teachR::make.sheet("[path to file]/ad_practical03.Rmd", "and")
library(english)
```


## Today

- Basic descriptive visualisation
- What makes a good plot
- Including plots in papers/reports
- Write-up: materials

## Guided part


First of all, let's make sure we're all set up:

- Log into Canvas and go to this week's practical
- Download the "prac_04_visual.Rmd" file into your analysing_data project folder
- Open RStudio

\ 

### Read-in and clean

We can use code from last practical to read in and clean our data. Note that the code below is a condensed version of what we did last week but it does the same job.

```{r}
library(tidyverse)
theme_set(theme_grey())
# read in
gensex <- read_csv("https://mivalek.github.io/adata/gen_sex_q.csv")


### AGE

# recode age and gender
gensex <- gensex %>%
  mutate(Age = recode(Age, "18 years" = "18", "19 years old" = "19"),
         Age = as.numeric(Age),
         Gender = factor(Gender))

# count how many cases you're about to remove
age_removed <- gensex %>% filter(Age > 99) %>% nrow()

# count how many you're about to remove
all_missing <- gensex %>% filter(rowSums(is.na(gensex)) > 10) %>% nrow()

# remove them
gensex <- gensex %>% filter(Age < 100 & rowSums(is.na(gensex)) < 11)


# DESCRIPTIVES ------------------------------------------------------------

# desc for Age

age_desc <- gensex %>%
  summarise(mean = mean(Age, na.rm = T),
            sd = sd(Age, na.rm = T),
            min = min(Age, na.rm = T),
            max = max(Age, na.rm = T)) %>%
  modify(round, 2)

# Break-down of gender with age stats
gen_age_desc <- gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>%
  group_by(Gender) %>%
  summarise(n = n(),
            perc = n()/nrow(gensex) * 100,
            mean_age = mean(Age, na.rm = T),
            sd_age = sd(Age, na.rm = T))


# # give column names (with markdown)
# # give table a caption
# # round to 2 decimal places
# gen_age_desc %>%
#   mutate(Gender = str_to_sentence(Gender)) %>%
#   kable(col.names = c("Gender", "*N*", "%", "*M*~age~", "*SD*~age~"),
#         caption = "Table 1 *Descriptive statistics by Gender*",
#         digits = 2) %>%
#   kable_styling()

```


### Visualisation

While some descriptive statistics and results are best summarised in a table, it is often worth visualising your data using plots. Some plots are great for presenting results to your audience while others are just a tool for you to get a better feel for your data and results. It is neither necessary nor desirable to include every plot you create in your write-up/paper/poster. Only show plots if they add anything of explanatory value to your body text and tables, the rest are for your eyes only (or for the appendix/supplementary materials).

Let's look at a simple histogram of our `Age` variable:

```{r, message=T}
gensex %>%
  ggplot(aes(x = Age)) +
  geom_histogram()
```

\ 

OK, this histogram gives us a visual representation of the distribution of ages in the sample: The vast majority of participants are 18-19 years old. This is not news as we can read this from the first summary we ran:

```{r}
gensex %>% summary()
```

```{r, echo = F}
q1 <- quantile(gensex$Age, .25)
q3 <- quantile(gensex$Age, .75)
```

As you can see the first quartile of `age` is `r q1` and the third quartile is `r q3`. That means that at least the middle 50% of the sample's age falls within `r q1` and `r q3`. Seeing the distribution plotted is, however, still useful.

Aesthetically speaking, our histogram isn't that much to look at. Fortunately, with `ggplot()` and some of its extensions, we can customise our plots to death and beyond! The `cowplot` package contains a nice clean theme for publication-style `ggplot()` plots. Let's install it.

```{r, eval=F}
install.packages("cowplot")
```

Once we load it, we can add `theme_cowplot()` to our plot to make it look a little better:

```{r, message=T}
library(cowplot)
gensex %>%
  ggplot(aes(x = Age)) +
  geom_histogram() +
  theme_cowplot()
```

Notice that `R` warns us that the histogram plots values by default into 30 bins. In other words, it leaves enough room in the plot for 30 columns. However, we only have 8 distinct ages (17-24). We might as well change the number of bins to 8 using the `bins=` argument inside `geom_histogram()`. While we are at it, we can also change the colour of the outline of the bars (`color=`) and the colour of the bars themselves (`fill=`):

```{r}
library(cowplot)
gensex %>%
  ggplot(aes(x = Age)) +
  geom_histogram(bins = 8, color = "black", fill = "grey") +
  theme_cowplot()
```

\ 

That looks way better! The only thing left to do now is give our axes proper labels:

```{r}
library(cowplot)
gensex %>%
  ggplot(aes(x = Age)) +
  geom_histogram(bins = 8, color = "black", fill = "grey") +
  labs(x = "Age in years", y = "Frequency") +
  theme_cowplot()
```

\ 

There, that's one neat plot!

\ 

OK, let's see what we can do with our `Gender` variable.

Basically the only way of visualising a single categorical variable is using the barchart:

```{r}
gensex %>%
  ggplot(aes(x = Gender)) +
  geom_bar()
```

Again, we can customise the plot in several ways. Notice that the bars are ordered alphabetically and their labels are the same as the levels of our `Gender` variable:

```{r}
gensex %>%
  pull(Gender) %>%
  levels()
```


\ 

Let's tweak the appearance of our plot using what we already know:

```{r}
gensex %>%
  ggplot(aes(x = Gender)) +
  geom_bar(color = "black", fill = "seagreen") +
  labs(x = "Gender", y = "N") +
  theme_cowplot()
```

<!-- \  -->

<!-- As you can see, we have some overlapping labels. That's not great. There are several ways around this issue. For instance, we could rotate the labels 45° to stop them from overlapping. However, I think the neatest way is simply to flip the plot on its side. The `coord_flip()` command does exactly that: -->

<!-- ```{r} -->
<!-- gensex %>% -->
<!--   mutate(Gender = str_to_sentence(Gender), -->
<!--          Gender = fct_infreq(Gender)) %>% -->
<!--   ggplot(aes(x = Gender)) + -->
<!--   geom_bar(color = "black", fill = "grey") + -->
<!--   labs(x = "Gender", y = "N") + -->
<!--   theme_cowplot() + -->
<!--   coord_flip() -->
<!-- ``` -->

I think this one is also rather nice!

\ 

<div class="solText">
If you really care about the small things and want to replace the `NA` with something more explicit, you can use the `fct_explicit_na()` function that turns `NA` into `(Missing)` by default (but this can be changed to whatever you want). Again, because we're essentially changing the variable, the command needs to be inside `mutate()`:

```{r}
gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>%
  ggplot(aes(x = Gender)) +
  geom_bar(color = "black", fill = "grey") +
  labs(x = "Gender", y = "N") +
  theme_cowplot()
```


*Note* that the `mutate()` command comes before the data are pushed into `ggplot()`. That is because we aren't really changing the plot itself but the `gensex` tibble. We are essentially telling `R` to temporarily modify the `Gender` variable in the tibble before it gets used for plotting.
</div>

\ 

Next, we can visualise the age breakdown by gender choice using boxplots. To do this, you need to map each of these variables onto one axis inside of the `aes()` function: The categorical variable belongs on the X axis and the continuous on the Y axis:

```{r}
gensex %>%
  ggplot(aes(x = Gender, y = Age)) +
  geom_boxplot()
```

\ 

We can re-use essentially all the commands from our previous plot to make this boxplot by group a little nicer.

```{r}
gensex %>%
  mutate(Gender = fct_explicit_na(Gender)) %>%
  ggplot(aes(x = Gender, y = Age)) +
  geom_boxplot(color = "black", fill = "grey") +
  labs(x = "Gender", y = "Age in years") +
  theme_cowplot()
```

\ 

```{r, echo=F, eval=F}
gensex %>%
  pivot_longer(-(1:3),
               names_to = "item",
               values_to = "response") %>%
  ggplot(aes(x = response)) +
  geom_bar() +
  facet_wrap(~item) +
  theme_cowplot()
```

That's it for descriptive visualisation. Let's build up a more interesting plot that will show us something about the relationships between the main measured variables.

Suppose we are interested in knowing whether there is a relationship between participants' self-ratings of masculinity and femininity. If these are inverses of each other as is sometimes claimed, the dots should be pretty much lined up: the more masculine you think you are, the less feminine you consider yourself. However, these dimensions may well not be opposites of one another. If that's the case, it is possible for a person to feel both highly masculine and feminine at the same time. Let's build a plot to find out.

Since both of these variables are continuous, a scatterplot is probably the best option. In `ggplot()`, we create scatterplots using `geom_point()`:

```{r}
gensex %>%
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1)) +
  geom_point() 
```

\ 

OK, there is definitely a moderate-to-strong negative relationship between a person's masculinity and femininity self-rating but it's hardly perfect. Some people are relatively high on both of these dimensions!

The plot also looks like we only have a handful of observations because the points are drawn on top of each other. We can jitter them around to get a better reflection of our sample size. Now, jittering points makes the plot lie a little about the actual values people chose but the trade-off is that the sample size shows better.

```{r}
gensex %>%
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1)) +
  geom_point(position = "jitter")
```  

\ 

That's better.

Now, let's say we want to know if the relationship looks the same for each level of the `Gender` variable. We can choose to colour the points by gender. **Since we're mapping a variable, we need to do this as an &aelig;sthetic using the `aes()` function**:

```{r}
gensex %>%
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1, colour = Gender)) +
  geom_point(position = "jitter")
```

Finally, we can change the look of the plot a little. Let's make the points a little larger using the `size=` argument and ever so slightly transparent (`alpha=` goes from 0 to 1). Then add better axis labels and change the theme of the plot.

```{r}
gensex %>%
  ggplot(aes(x = Gender_fem_1, y = Gender_masc_1, colour = Gender)) +
  geom_point(position = "jitter", size = 2, alpha = .4) +
  labs(x = "Femininity", y = "Masculinity") +
  theme_cowplot()
```

That's a nice one!

It's interesting to see that different gender categories display different patterns. While participants who reported their gender as female seem to be roughly spread across the diagonal of the plot, those who reported their gender as male seem to be distributed more flatly in the top half of the plot. Unfortunately, we do not have enough data from participants in the "Other" category to make any observations.

Please bear in mind that we cannot really say for certain if these differences are statistically significant just by looking at the scatterplot. For that, we need to run some tests!

\ 

This is just about enough plotting for one day. If you've been typing along, save your script.


<font size="+2">**Over to you!**</font>

\ 


## Worksheet

```{r, echo = F}
data <- read_csv("https://raw.githubusercontent.com/mivalek/AnD/master/data/harm_data.csv")
```

In the remaining time you will practice visualising data with `ggplot()`. We will be using the same data from the [paper by Swiderska and K&uuml;ster (2018)](https://journals.sagepub.com/doi/10.1177/0301006618809919) as we did last time.

`r task()`First of all, if you haven't finished last week's tasks pick up where you left in the [week 3 practical worksheet](https://mivalek.github.io/adata/prac/prac_03_wkst.html#worksheet). There is a [version with solutions](https://mivalek.github.io/adata/prac/sol/prac_03_wkst_sol.html#worksheet) available now if you need help but when you use it, make sure you understand the code! There's no point copy-pasting the solutions just for the sake of completing the tasks.

`r task()`Download the "prac_04_visual.Rmd" file from Canvas and open it in RStudio. It is an expanded version of the one you were working on last week. Practice makes perfect...

Use the R Markdown file to complete the following tasks.

`r task()`Complete the `packages`, `read-in`, `inspect`, and `descriptives` code chunks, as per last week's tasks. You have done it before so it would be great if you could do the tasks without further prompting. If you need a little help, that's OK. Just check out the worksheet for the individual tasks. If you must, you can even check the code you used last time but try not to copypaste it here. Rather, type it in manually to help develop coding habits.

```{r, eval=F, echo= solution, sol=T}
# packages chunk

library(tidyverse)
library(cowplot)
# if kableExtra doesn't work
library(knitr)
# otherwise
library(kableExtra)
```


```{r, echo= solution, sol=T}
# read-in chunk

data <- read_csv("https://raw.githubusercontent.com/mivalek/AnD/master/data/harm_data.csv")
```


```{r, echo= solution, sol=T}
# inspect chunk

data <- data %>%
  filter(Age != "1y") %>% # filter out row with typo in Age
  mutate(Age = as.numeric(Age), # turn into numeric
         # turn nominal variables into factors, giving them labels
         ID = factor(ID),
         Condition = factor(Condition, labels = c("Human_Unharmed", "Human_Harmed",
                                                  "Robot_Unharmed", "Robot_Harmed")),
         Humanness = factor(Humanness, labels = c("Human", "Robot")),
         Harm = factor(Harm, labels = c("Unharmed", "Harmed")),
         Gender = factor(Gender, labels = c("Male", "Female")))

# Number of ppts we're about to remove due to outlying/unlikely ages
too_young <- data %>%
  filter(Age < 17) %>%
  nrow()
too_old <- data %>%
  filter(Age > 90) %>%
  nrow()

# remove cases
data <- data %>% filter(Age > 18 & Age < 90)
```


```{r, echo=solution, class = "solution"}
# descriptives chunk

# descriptives of Age
age_desc <- data %>%
  summarise(mean = mean(Age),
            sd = sd(Age),
            min = min(Age),
            max = max(Age)) %>%
  modify(round, 2)

# Ns + Age breakdown by gender
gender_desc <- data %>%
  mutate(Gender = fct_explicit_na(Gender)) %>%
  group_by(Gender) %>%
  summarise(n = n(),
            perc = n()/nrow(data) * 100,
            age_mean = mean(Age),
            age_sd = sd(Age))

# descriptives of measured variables by type
type_desc <- data %>%
  group_by(Humanness) %>%
  summarise(n = n(),
            pain_m = mean(Pain),
            exp_m = mean(Experience),
            agency_m = mean(Agency),
            consc_m = mean(Consciousness))

# ... and by type for Unharmed avatars only
unharm_desc <- data %>%
  filter(Harm == "Unharmed") %>%
  group_by(Humanness) %>%
  summarise(n = n(),
            attr_m = mean(Attractiveness),
            emp_m = mean(Empathy))
```


`r task()`Complete the code for the histogram of age, bar chart of gender and boxplot of age by gender in the `plots` chunk. They should look like this (feel free to choose the colours you like):

```{r, include=solution, sol=T}
# you can store plots iside objects too!
age_hist <- data %>%
  ggplot(aes(x = Age)) +
  geom_histogram(fill = "white", colour = "black") +
  labs(x = "Participants' age in years", y = "N") +
  theme_cowplot()

gender_bar <- data %>%
  ggplot(aes(x = Gender)) +
  geom_bar(fill = "seagreen") +
  labs(x = "Participants' gender", y = "N") +
  theme_cowplot()

gender_age_box <- data %>%
  ggplot(aes(x = Gender, y = Age)) +
  geom_boxplot()+
  labs(x = "Gender", y = "Participants' age in years") +
  theme_cowplot()
```  
  
```{r, echo=F, fig.width=9, fig.height=3}
plot_grid(age_hist, gender_bar, gender_age_box, nrow = 1)
```

<div class="solText">
Since the plots are assigned to objects using the `<-` operator, they will not be shown when you run the code in the chunk. To see what they look like you need to either only run the part of the code to the right of the `<-` or run the chunk and then type the name of the object into the console.</div>

`r task()`Complete the code for the `pain_consc` plot so that it looks like this:

```{r, include=solution, sol=T}
pain_consc <- data %>%
  ggplot(aes(x = Pain, y = Consciousness, colour = Harm)) +
  geom_point(position = "jitter") +
  facet_wrap(~Humanness) +
  theme_bw()
```

```{r, echo=F}
pain_consc
```

Notice how the `facet_wrap()` function is used to create two plots faceted by the `Humanness` variable.


`r task()`Complete the **Write-up** section of the .Rmd file.

<!-- solution
<pre>
<code>## Method

### Participants

The study was conducted on a sample of &#96;r nrow(data)&#96; volunteers (*M*~age~ = &#96;r age_desc %>% pull(mean)&#96;, *SD*~age~ = &#96;r age_desc %>% pull(sd)&#96;). The data were collected anonymously on-line. Data from &#96;r too_young + too_old&#96; participants were excluded due to unlikely values of age. Table 1 shows the distribution of gender as well as an age brake-up by individual gender categories.

### Materials

We used images of faces of a photorealistic human avatar and a robotic avatar as stimuli. These two faces were then modified to create the faces’ harmed analogues by adding the facial texture of a moderate burn to their right side. All images measured 960&times;720 pixels and were displayed individually on a white background.

We measured the extent to which participants attributed the capacity to experience (7 items), agency (7 items), consciousness (2 items), and pain (1 item) to the avatars. A 7-item measure of empathy with avatars was used and avatars were also rated on attractiveness (1 item). All items used a 7-point likers scale.

### Procedure

Participants were presented with pictures of the avatars. Their task was to evaluate the degree to which mental capacities (experience, agency, consciousness, and pain) could be attributed to the faces and the extent to which the presented avatars elicited empathic reactions. Every page of the survey consisted of the respective face displayed above a 7-point, Likert-type response scale (1 = "Strongly disagree" to 7 = "Strongly agree"). The survey was delivered via EFS Survey (Version 9.0, QuestBack AG, Germany). The experiment followed a 2 (Harm: harmed vs. control) 2&times;2 (Robotization: human vs. robotic) between-subjects factorial design.</code>
</pre>

\ 

The R Markdown above will appear as:

#### 4.1 Method

##### 4.1.1 Participants

The study was conducted on a sample of `r nrow(data)` volunteers (*M*~age~ = `r age_desc %>% pull(mean)`, *SD*~age~ = `r age_desc %>% pull(sd)`). The data were collected anonymously on-line. Data from `r too_young + too_old` participants were excluded due to unlikely values of age. Table 1 shows the distribution of gender as well as an age brake-up by individual gender categories.

##### 4.1.2 Materials

We used images of faces of a photorealistic human avatar and a robotic avatar as stimuli. These two faces were then modified to create the faces’ harmed analogues by adding the facial texture of a moderate burn to their right side. All images measured 960&times;720 pixels and were displayed individually on a white background.

We measured the extent to which participants attributed the capacity to experience (7 items), agency (7 items), consciousness (2 items), and pain (1 item) to the avatars. A 7-item measure of empathy with avatars was used and avatars were also rated on attractiveness (1 item). All items used a 7-point likers scale.

##### 4.1.3 Procedure

Participants were presented with pictures of the avatars. Their task was to evaluate the degree to which mental capacities (experience, agency, consciousness, and pain) could be attributed to the faces and the extent to which the presented avatars elicited empathic reactions. Every page of the survey consisted of the respective face displayed above a 7-point, Likert-type response scale (1 = "Strongly disagree" to 7 = "Strongly agree"). The survey was delivered via EFS Survey (Version 9.0, QuestBack AG, Germany). The experiment followed a 2 (Harm: harmed vs. control) 2&times;2 (Robotization: human vs. robotic) between-subjects factorial design.

<div class="warn">
**We lifted some of the above from the original paper. You should  absolutely not do that!**
</div>-->

`r task()`Edit the code in the `table_1` code chunk, giving it your tibble with age breakdown by gender, to create a nice formatted table in your document. Make sure the table show values to 2 decimal digits and to reference your table somewhere in the body text.


```{r table_1, echo = solution, sol=T}
gender_desc %>%
  kable(col.names = c("Gender", "*N*", "%", "*M*~age~", "*SD*~Age~"),
        caption = "Table 1 *Descriptive statistics by categories of gender*",
        digits = 2) %>%
  kable_styling()
```


`r task()`Use the `fig_1` code chunk to print the grouped scatterplot. Notice how we specified the figure caption. All figures in a paper/report should have captions.

**Hint: **Just put the name of the object that contains the plot inside the chunk.

```{r, eval = F, echo = solution, sol=T, fig.cap="Figure 1. Relationship between pain and consciousness rating by levels of harm for human and non-human avatars."}
pain_consc
```

`r task()`Inspect the scatterplot and think about the patterns you see. Talk to your classmate about what the relationships shown in the plot might be.

`r task()`Use the `append_figs` code chunk to print the descriptive plots you made earlier in the Appendix.

Descriptive plots are rarely important enough to warrant space in the results section but they can be useful for the reader and so can be included in the Appendix/Supplementary materials.

```{r, eval = F, echo = solution, sol=T}
age_hist

gender_bar

gender_age_box
```

`r task()`Knit (generate) the final document from your R markdown file and rejoice in its beauty.

\ 

This week you learnt how to make pretty data visualisations and you added a Materials section to your write-up. Next time, we will review everything we've learned so far and practice the skills needed to prepare data for analysis and hypotheses testing, as you'll do in your lab report. But, for now...

<font size="+4">**Well done!**</font>

\ 

\ 



